---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r load-data}
act<-read.csv("activity.csv")

```


## What is mean total number of steps taken per day?
The total number of steps per day is
```{r nsteps}
stepsDay<-aggregate(act[,1],list(act$date),sum,na.rm=TRUE)
stepsDay
```
Histogram
```{r plot1}
hist(stepsDay$x)
```

The mean and median per day
```{r stats}
library(plyr)
stats<-ddply(act,~date,summarise,mean=mean(steps,na.rm=TRUE),median=median(steps,na.rm=TRUE))
stats
```


## What is the average daily activity pattern?
```{r plot2}
stepsInt<-aggregate(act[,1],list(act$interval),sum,na.rm=TRUE)
plot(stepsInt$Group.1,stepsInt$x,type="l",xlab="Interval",ylab="Total Steps")

```
```{r, results='hide'}
maxStInt<-stepsInt[stepsInt$x==max(stepsInt$x),1]
```
The  5-minute interval that contains the maximum number of steps is `r maxStInt`



## Imputing missing values

```{r}
missVal<-sum(is.na(act$steps))

```
The total number of missing values is `r missVal`.

A new data set will contain the same data with NAs values changed by the media in the interval 

```{r}
stepsIntMean<-aggregate(act[,1],list(act$interval),mean,na.rm=TRUE)
act2<-act
for (i in 1:length(act$steps)){
        if (is.na(act2[i,1])==TRUE){
                act2[i,1]<-stepsIntMean[stepsIntMean$Group.1==act2[i,3],2]
        }
}
```

Histogram of the new data set
```{r plot3}
stepsDay2<-aggregate(act2[,1],list(act2$date),sum,na.rm=TRUE)
hist(stepsDay2$x)
```

The mean and median per day
```{r stats2}
stats2<-ddply(act2,~date,summarise,mean=mean(steps,na.rm=TRUE),median=median(steps,na.rm=TRUE))
stats2
```

The correlation plot between the mean of the two data sets (with and without NAs) shows that the results are similar
```{r plot4}
plot(stats$mean,stats2$mean)
```

The values for the median are similar for the days in which there are not NAs (median=0) and differs in the days in which all steps were NA (median=34.11321)



## Are there differences in activity patterns between weekdays and weekends?

 First create the data set with a new column
 
```{r data3}
act3<-act2
act3$Day<-weekdays(as.Date(act2$date))
act3[act3$Day=="domingo"|act3$Day=="sábado",]$Day<-"weekend"
act3[act3$Day=="lunes"|act3$Day=="martes"|act3$Day=="miércoles"|act3$Day=="jueves"|act3$Day=="viernes",]$Day<-"weekday"
act3$Day<-as.factor(act3$Day)
```

Create data sets for weekdays and weekends an plot the average of steps by interval

```{r plot5}
Weekend<-act3[act3$Day=="weekend",]
Weekday<-act3[act3$Day=="weekday",]
stepsIntWD<-aggregate(Weekday[,1],list(Weekday$interval),mean)
stepsIntW<-aggregate(Weekend[,1],list(Weekend$interval),mean)
par(mfrow=(c(2,1)))
plot(stepsIntW$Group.1,stepsIntW$x,type="l",main="Weekend",ylab="Steps",xlab="Interval")
plot(stepsIntWD$Group.1,stepsIntWD$x,type="l",main="Weekday",ylab="Steps",xlab="Interval")
```

There are significant differences in the activity depending on wether the day is weekday or weekend. 

